Creating the ``r params$package_name`` R package

### 2025-10-07 create-FSHybridPLS.Rmd
curve_normalize

We start by specifying the information needed in the DESCRIPTION file of the R package.

```{r package-setup, message=FALSE, results='hide'}
usethis::create_package(
  path = ".",
  fields = list(
    Package = params$package_name,
    Version = "0.0.0.9000",
    Title = "A Package That Says Hello",
    Description = "This package says hello.  But its actual purpose is to show how an R package can be completely coded in a single R markdown file.",
    Imports = "fda (>= 6.1.3), Matrix, funData, foreach, irlba, refund",
    `Authors@R` = person(
      given = "First",
      family = "Last",
      email = "you@gmail.com",
      role = c("aut", "cre")
      )
  )
)
usethis::use_mit_license(copyright_holder = "F. Last")
```

@@

@@

```{r}
normailize.all <- function(split.all_obj){
  curve_normalize_result_1 <- curve_normalize_train_test_fd(
    split.all_obj$data_fun_1_train,
    split.all_obj$data_fun_1_test)
  data_fun_1_train <- curve_normalize_result_1$train
  data_fun_1_test <- curve_normalize_result_1$test

@@

>>>>

We start by specifying the information needed in the DESCRIPTION file of the R package.
 library()
library(fda)
```{r package-setup, message=FALSE, results='hide'}
usethis::create_package(
  path = ".",
  fields = list(
    Package = params$package_name,
    Version = "0.0.0.9000",
    Title = "A Package That Says Hello",
    Description = "This package says hello.  But its actual purpose is to show how an R package can be completely coded in a single R markdown file.",
    Imports = "fda (>= 6.1.3), Matrix, funData, foreach, irlba, refund, R6",
    `Authors@R` = person(
      given = "First",
      family = "Last",
      email = "you@gmail.com",
      role = c("aut", "cre")
      )
  )
)
usethis::use_mit_license(copyright_holder = "F. Last")
```

@@

```{r}
#' @title Generate Train Indices
#' @description Generates train/test split indices.
#' @param n_sample Integer. Number of samples.
#' @param train_ratio Numeric. Ratio of train samples.
#' @return Vector of train indices.
get_idx_train <- function(n_sample, train_ratio) {
  n_train <- floor(n_sample * train_ratio)
  idx_train <- sort(sample(n_sample, n_train, replace = FALSE))
  return(idx_train)
}
```

@@

### split.all

Splits functional predictors, scalar predictors, and response into train and test sets.

```{r}
#'Split Functional, Scalar, and Response Data
#' @param data_fun_1 fd object (fda package). Functional predictor 1.
#' @param data_fun_2 fd object (fda package). Functional predictor 2.
#' @param data_scalar Matrix. Scalar predictors.
#' @param response Vector. Response variable.
#' @param train_idx Vector. Indices for train data.
#' @return List containing split train and test sets.
#' @export
split.all <- function(data_fun_1, data_fun_2, data_scalar, response, train_idx) {
  # --- Type Checks ---
  stopifnot(
    # Functional data checks (assuming 'fd' class from the 'fda' package)
    inherits(data_fun_1, "fd"),
    inherits(data_fun_2, "fd"),

    # Scalar predictor check
    is.matrix(data_scalar),

    # Response variable check
    is.vector(response) || is.numeric(response) || is.factor(response),

    # Training index check
    is.vector(train_idx),
    is.numeric(train_idx),
    all(train_idx > 0), # Indices should be positive
    all(train_idx == round(train_idx)) # Indices should be integers
  )
  # -------------------

  return(list(
    data_fun_1_train = data_fun_1[train_idx],
    data_fun_1_test = data_fun_1[-train_idx],
    data_fun_2_train = data_fun_2[train_idx],
    data_fun_2_test = data_fun_2[-train_idx],
    data_scalar_train = data_scalar[train_idx, , drop = FALSE],
    data_scalar_test = data_scalar[-train_idx, , drop = FALSE],
    response_train = response[train_idx],
    response_test = response[-train_idx]
  ))
}
```

### normailize.all

```{r}
#' 
#' @export
normailize.all <- function(split.all_obj){
  curve_normalize_result_1 <- curve_normalize_train_test_fd(
    split.all_obj$data_fun_1_train,
    split.all_obj$data_fun_1_test)
  data_fun_1_train <- curve_normalize_result_1$train
  data_fun_1_test <- curve_normalize_result_1$test

@@

### curve_normalize
Subtracts the mean function and divides by a scaling constant (deno). This implementation uses the native fda S3 methods (minus.fd and times.fd).
```{r}
#' Normalize a functional data object by its mean and standard deviation.
#'
#' @param functional fd object (fda package). The functional data to be normalized (N samples).
#' @param mean_functional fd object (fda package). The mean function (1 sample).
#' @param deno Numeric scalar. The standard deviation or scaling factor.
#' @return A new fd object containing the normalized functions.
#' @export
curve_normalize <- function(functional, mean_functional, deno) {
  # Type checks
  stopifnot(
    "'functional' must be an 'fd' object" = inherits(functional, "fd"),
    "'mean_functional' must be an 'fd' object" = inherits(mean_functional, "fd"),
    "'deno' must be a single positive numeric value" = is.numeric(deno) && length(deno) == 1 && deno > 0
  )

  # 1. Centering: functional - mean_functional
  # fda::minus.fd handles the broadcasting of the single-sample mean_functional.
  centered_functional <- fda::minus.fd(functional, mean_functional)

  # 2. Scaling: (1/deno) * centered_functional
  functional_normalized <- fda::times.fd(1/deno, centered_functional)

  return(functional_normalized)
}
```

```{r}
#' 
#' @export
curve_normalize <- function(functional, mean_functional, deno){
  # operation also takes care of the original_X

  functional_normalized <- fd(coef =
                                coef(functional) - coef(mean_functional) %*% matrix(1, ncol = ncol(coef(functional))),
                              basisobj = functional$basis
                              )#numerator
  functional_normalized <- times.fd(1/deno, functional_normalized)
  return(functional_normalized)
}
```

```{r}
curve_normalize_train_test <- function(train, test){
  train_normalized <- train
  test_normalized <- test
  deno_train <- rep(NA, train@n_functional)
  for (k in 1:train@n_functional){ # for kth functional predictor
    mean_train <- mean.fd(train@functional_list[[k]]) # train mean function
    centered_train <- center.fd(train@functional_list[[k]])
    sd_train <- sd.fd(train@functional_list[[k]])
    deno_train[k] <- sqrt(inprod(sd_train, sd_train))

    train_normalized@functional_list[[k]] <- curve_normalize(
      train_normalized@functional_list[[k]],
      mean_train,
      deno_train[k]
      )

    test_normalized@functional_list[[k]] <-
      curve_normalize(
        test_normalized@functional_list[[k]],
        mean_train,
        deno_train[k]
      )
  }
  return( list(
    train = train_normalized,
    test = test_normalized,
    deno_train = deno_train
    ))}
```

```{r}
# scalar_normalize <- function(scalar_predictor, mean, deno){
#   scalar_predictor_normalized <- subtr_broadcast(scalar_predictor, mean) #numerator
#   deno_mat <- matrix(rep(1,nrow(scalar_predictor))) %*% deno
#   scalar_predictor_normalized <- scalar_predictor_normalized / deno_mat
#   return(scalar_predictor_normalized)
# }
#
# scalar_normalize_train_test <- function(predictor_train, predictor_test){
#   predictor_train_normalized <- predictor_train
#   predictor_test_normalized <- predictor_test
#
#   mean_train <- get_mean(predictor_train@Z) # train mean function
#   sd_train <- get_sd(predictor_train@Z)
#
#   predictor_train_normalized@Z <-
#     scalar_normalize(
#       predictor_train@Z,
#       mean_train,
#       sd_train
#       )
#
#   predictor_test_normalized@Z <-
#     scalar_normalize(
#       predictor_test@Z,
#       mean_train,
#       sd_train
#     )
#
#   return(
#     list(
#       train = predictor_train_normalized,
#       test = predictor_test_normalized
#     )
#   )
# }
```

```{r}
#
# btwn_normalize_train_test <- function(train, test){
#   #"Scale the scalar predictors so that the variability between the functional and scalar predictors are comparable:")
#   Z_s <- train@Z
#   train_normalized <- train
#   test_normalized <- test
#
#   omega <- 0
#   for (k in 1:train@n_functional){ # for kth functional predictor
#     omega <- omega + sum(diag(inprod(
#       train@functional_list[[k]],
#       train@functional_list[[k]]
#       )))
#
#   }
#     omega <- omega / sum((Z_s)^2)
#
#     train_normalized@Z <- sqrt(omega) * (train@Z)
#     test_normalized@Z <- sqrt(omega) * (test@Z)
#
#     return(
#       list(
#         train = train_normalized,
#         test = test_normalized
#       ))}
```

### DataSplitterNormalizer
R6 class to split functional and scalar predictors into train and test sets and apply normalization techniques including curve normalization, scalar normalization, and between-functional-scalar normalization.
```{r}
#' @title Dataset Splitting and Normalization
#' @description 
DataSplitterNormalizer <- R6::R6Class(
  "DataSplitterNormalizer",

  public = list(

    data_original = NULL,  # Stores the original dataset
    n_sample = NULL,  # Number of samples
    train_ratio = NULL,  # Train-test split ratio
    idx_train = NULL,  # Train indices
    train=NULL,
    test=NULL,

    #' @title Initialize Data Splitter and Normalizer
    #' @description Constructs the instance, splits the data into train and test sets.
    #' Users must call `normalize_data()` separately.
    #' @param data_original List. Contains `functional_predictors` (list of `fd` objects)
    #' and `scalar_predictors` (matrix).
    #' @param train_ratio Numeric. Ratio of data to use for training (e.g., 0.8).
    #' @return An instance of `DataSplitterNormalizer`.
    #' @export
    initialize = function(data_original, train_ratio = 0.8) {
      self$n_sample <- length(data_original$response)
      self$train_ratio <- train_ratio
      self$data_original <- data_original

    },

    #' @title Split Data into Train and Test Sets
    #' @description Creates train/test indices and splits functional, scalar, and response variables.
    #' @return None. Updates train and test variables.
    #' @export
    split_data = function(idx_train = NULL) {
      # Use provided idx_train if available; otherwise, generate it
      if (is.null(idx_train)) {
        self$idx_train <- get_idx_train(self$n_sample, self$train_ratio)
      } else {
        self$idx_train <- idx_train
      }

      split_packed <- split.all(
        self$data_original$functional_1,
        self$data_original$functional_2,
        self$data_original$scalar,
        self$data_original$response,
        self$idx_train
      )

      # Assign train/test sets
      self$train <- list(
        functional_1 = split_packed$data_fun_1_train,
        functional_2 = split_packed$data_fun_2_train,
        scalar = split_packed$data_scalar_train,
        response = split_packed$response_train
      )
      self$test <- list(
        functional_1 = split_packed$data_fun_1_test,
        functional_2 = split_packed$data_fun_2_test,
        scalar = split_packed$data_scalar_test,
        response = split_packed$response_test
      )

    },

    #' @title Normalize Train and Test Data
    #' @description Applies curve, scalar, and between-functional-scalar normalization.
    #' @return None. Updates train and test sets.
    #' @export
    normalize_data = function() {
      # Initialize normalizers
      normalizer_functional <- normalizer_functional_predictor$new()
      normalizer_scalar <- normalizer_scalar_predictor$new()

      # Normalize first functional predictor
      normalizer_functional$normalize(self$train$functional_1, self$test$functional_1)
      self$train$functional_1 <- normalizer_functional$get_normalized()$train_normalized
      self$test$functional_1 <- normalizer_functional$get_normalized()$test_normalized

      # Normalize second functional predictor
      normalizer_functional$normalize(self$train$functional_2, self$test$functional_2)
      self$train$functional_2 <- normalizer_functional$get_normalized()$train_normalized
      self$test$functional_2 <- normalizer_functional$get_normalized()$test_normalized

      # Normalize scalar predictors
      normalizer_scalar$normalize(self$train$scalar, self$test$scalar)
      self$train$scalar <- normalizer_scalar$get_normalized()$train_normalized
      self$test$scalar<- normalizer_scalar$get_normalized()$test_normalized

      # Apply between-functional-scalar normalization
      self$normalize_between()

      # Normalize response variable
      self$normalize_response()
    },

    #' @title Normalize Between Functional and Scalar Predictors
    #' @description Scales scalar predictors to ensure comparability with functional predictors.
    #' @export
    normalize_between = function() {
      omega <- sum(diag(inprod(self$train$functional_1, self$train$functional_1))) +
        sum(diag(inprod(self$train$functional_2, self$train$functional_2)))

      omega <- omega / sum(self$train$scalar^2)

      self$train$scalar <- sqrt(omega) * self$train$scalar
      self$test$scalar <- sqrt(omega) * self$test$scalar
    },

    #' @title Normalize Response Variable
    #' @description Centers response variable based on training mean.
    #' @export
    normalize_response = function() {
      response_mean_train <- mean(self$train$response)
      self$train$response <- self$train$response - response_mean_train
      self$test$response <- self$test$response - response_mean_train
    } #must not end with ,

  )
)
